---
title: "Miniproject3"
author: "Dianne Caravela, Neyda Carrillo"
date: "4/16/2019"
output:   
  html_document:
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(macleish)
library(leaflet)
library(tidyverse)
library(sf)
library(rgdal)
library(plotly)
```

# Wellness at Macleish: It’s in Your Hands (and Feet)

500 to 800 words including relevant external information in the analysis and reporting at least one actionable recommendation (what should Smith DO with your findings?)

a short discussion on the resources necessary to implement your recommendation:
what will it cost?
how long will it take?
are there any concerns about environmental impact?
do you see any major roadblocks?

Link to GitHub^[https://github.com/dcaravela/miniproject3]

### Getting trail and elevation data
```{r, warning=FALSE, message=FALSE}
macleish_trails <- macleish_layers %>%
  pluck("trails") %>%
  mutate(length = st_length(geometry))

macleish_elevation <- macleish_layers %>%
  pluck("contours_3m")

```

### Joins trails and elevation data, creates new variables
```{r, warning=FALSE, message=FALSE, echo = TRUE}
trail_elevations <- st_join(macleish_trails, macleish_elevation) %>%
  group_by(name) %>%
  mutate(
    #finding change in elevation for each segment
    min_elev = min(ELEV_FT), 
          max_elev = max(ELEV_FT), 
          change_elev = max_elev - min_elev, 
    
    #convert to miles
          length_miles = length / 1609.344,
    
    #numeric difficulty rating based on Shenandoah's Hiking Difficulty
         difficulty_num = sqrt(change_elev * 2 * length_miles),
    
    #Rating of trail relative to the other Macleish trails
         difficulty_rating = case_when(difficulty_num <= 4 ~ "easy",
                                       difficulty_num >= 10 ~ "difficult",
                                       TRUE ~ "medium"))  %>%
  na.omit()
trail_elevations
```

### Creates leaflet graph
```{r, message=FALSE}
#function to map difficulty to the color palatte
rating <- colorNumeric(
  palette = "inferno",  
  domain = trail_elevations$difficulty_num)

#maps spacial data with leaflet
trail_map <- leaflet(data = trail_elevations) %>%
  addTiles() %>%
  
  #adds individual markers for each trail, showing trail name and difficulty info
  addMarkers(lng = -72.68025, lat = 42.45659168, popup = "East Loop Trail Medium (9.0)") %>%
  addMarkers(lng = -72.68037, lat = 42.454, popup = "Poplar Hill Road Difficult (11.0)") %>%
  addMarkers(lng = -72.679, lat = 42.4513, popup = "East Loop Trail Difficult (10.4)") %>%
  addMarkers(lng = -72.6795, lat = 42.4489, popup = "Entry Trail Easy (2.2)") %>%
  addMarkers(lng = -72.68085, lat = 42.4482, popup = "Poplar Hill Road Medium (4.4)") %>%
  addPolylines(color = ~rating(difficulty_num)) %>%
  addMarkers(lng = -72.6818, lat = 42.4490, popup = "West Loop Trail Medium (5.7)") %>%
  addMarkers(lng = -72.6834, lat = 42.4525, popup = "Porcupine Trail Medium (9.2)") %>%
  addMarkers(lng = -72.6837, lat = 42.4490, popup = "Vernal Pool Loop Trail Easy (3.0)") %>%
  addMarkers(lng = -72.6840, lat = 42.4555, popup = "West Loop Trail DIfficult (12.9)") %>%
  
  #adds legend at bottom of screen
  addLegend("bottomleft", pal = rating, values = ~difficulty_num, title = "Difficulty Rating", opacity = 1)

trail_map
```

